Planar Maximal Covering with Ellipses

### 2019-05-05 quali/quali.presentation/slides.tex
WIP

\usepackage[
backend=biber,style=authoryear,autocite=inline
]{biblatex}
%\DeclareLanguageMapping{english}{english-apa}

@@

\section{Introduction}
\begin{frame}
\frametitle{Introduction}
\begin{itemize}
	\item Covering problems
	\begin{itemize}
		\item Minimum Cover Problem
		\item Maximal Covering Problem \autocite{church:1974}
	\end{itemize}
	\item Maximal Covering Location Problem (MCLP)
	\item Planar Maximal Covering Location Problem (PMCLP)
	\begin{itemize}
	\item One disk: $\bigO(n^2)$ and $\bigO(n^2\log{n})$ algorithms
	\item $m$ disks: $\bigO(n^{2m-1}\log{n})$ algorithm
	\end{itemize}
	\item Goals
	\begin{itemize}
		\item Develop a $\bigO(n^2\log{n})$ algorithm for the one disk case
		\item Adapt it for the $m$ ellipses case creating a $\bigO(n^{2m})$ algorithm
	\end{itemize}
\end{itemize}

@@

\begin{frame}
	\begin{figure}
		\only<1>{\caption{Maximal cover by two disks.}}
		\only<2>{\caption{Minimum number of disks needed to cover the set of points.}}
		\only<1>{\input{figures/intro_a.tex}}
		\only<2>{\input{figures/intro_b.tex}}
		\source{Elaborated by the author.}
	\end{figure}
\end{frame}

@@

	$MCD(\Pp, 1)$ is the problem of placing one disk on the plane to cover a subset of a demand set $\Pp$ maximizing the weights of the covered points.

@@

	\begin{itemize}
		\item $\Pp=\{p_1,\dots,p_n\}$ is the demand set with weights $w(p_i)>0$
		\item $w(A)$, $A\subset \Pp$, is the sum of weights of the points in $A$
		\item $D(q)$ is a unit disk with center at point $q$
		\item \autocite{chazelle:1986} proposed a $\bigO(n^2)$ algorithm
		\item \autocite{drezner} proposed a $\bigO(n^2\log{n})$ which our work is based on
		\item We will introduce an equivalent problem...
	\end{itemize}
\end{frame}

@@

	\begin{itemize}
		\item The disks are fixed with centers at $\Pp =\{p_1,\dots,p_n\}$ with $w_k=w(p_k)$
		\item A clique is a non-empty intersection area of a subset of disks. We search for only a point in an optimal clique.
		\item An optimal solution for the maximum weight clique is an optimal solution for $MCD(\Pp,1)$.
	\end{itemize}
\end{frame}

\begin{frame}{Maximum Weight Clique Problem}{Equivalence}
		\begin{figure}[H]
		\centering

		\only<1>{\caption{An instance of the Maximum Weight Clique Problem obtained from an instance of $MCD(\Pp, 1)$.}}

		\only<2>{\caption{An instance of the Maximum Weight Clique Problem obtained from an instance of $MCD(\Pp, 1)$. In gray, the optimal solution.}}

		\only<1>{\input{figures/max_w_clique.tex}}
		\only<2>{\input{figures/max_w_clique_b.tex}}
      %\center{\includegraphics[width=.7\textwidth]{figures/mcd_instance.png}}
		\source{Elaborated by the author.}
		\label{fig:mcd_instance}
	\end{figure}
\end{frame}

\begin{frame}{Maximum Weight Clique Problem}{Algorithm}
	Defining $\Gamma_+(i,j)$ and $\Gamma_-(i,j)$:

	Let $D_i$ and $D_j$ be two unit disks that intersect at two points.

	\begin{itemize}
		\item Assume $D_i$ is at the origin
		\item We can find the polar angle of the intersection points. Let's call them $\Gamma_+(i,j)$ and $\Gamma_-(i,j)$ with the following condition:
		\begin{itemize}
			\item The arc defined by $\Gamma_+(i,j)$ and $\Gamma_-(i,j)$ bounds the region $D_i \cap D_j$

		\end{itemize}
		\item Later it will be shown how to find which one is which algorithmically.
	\end{itemize}
\end{frame}

@@

	\begin{itemize}
		\item There are $\bigO(n^2)$ intersections among $n$ disks

		\item Sorting takes $\bigO(n^2\log{n})$

@@

		\item It can be implemented in $K\log{n}$ where $K$ is the number of intersections.
	\end{itemize}

@@

		with $\le$ representing the set of covered points
	\end{block}

@@

	\begin{equation*}
	\max_q |\Pp \cap (q)|,
	\end{equation*}

	\begin{itemize}
		\item $E(q)$ is an axis-parallel ellipse with center point $q$
		\item Assuming unit weights for now
		\item Same algorithm for one disk
	\end{itemize}

@@

	\begin{itemize}
		\item $\E$ is a set of $m$ ellipses
		\item \autocite{canbolat} is the very first study on the problem. Slow exact method, a heuristic one was proposed.
		\item \autocite{andreta} proposed a method that breaks the problem into smaller optimization ones. Also, they developed a method for the non-axis-parallel case.
	\end{itemize}
\end{frame}

\begin{frame}[fragile]{Maximal Covering by Ellipses}{Pre-processing that finds every possible coverage for ellipse $E_i$}
	\begin{algorithm}[H]
		%\caption{Algorithm for $MCE(\Pp, a, b)$ with unit weights}\label{algoritmo:mce1}
		\begin{algorithmic}[1]
			\STATE $A \gets \bigcup_{j \in I_i} \{\Gamma_+(i,j) \cup \Gamma_-(i,j)\}$
			\STATE $Z \gets \{\}$

			\FOR{$cnt=1..2$} 
			\FOR{$a \in A$}
			\STATE Let $p_a$ be the point that intersects $E_i$ at angle $a$. 
			\IF{$a$ is a starting angle}
			\STATE $Cov \gets Cov \cup \{p_a\}$
			\ELSE
			\STATE $Cov \gets Cov \setminus \{p_a\}$
			\ENDIF
			\STATE $Z \gets Z \cup \{Cov\}$
			\ENDFOR
			\ENDFOR

@@

	\begin{itemize}
		\item The algorithm for $m$ ellipses tries every possible assignment of coverage for every one of the ellipses
		\item Run-time complexity of $\bigO(n^2)^m = \bigO(n^{2m})$

@@

		\item The unit-weight assumption can be easily dropped
	\end{itemize}

>>>>

\usepackage[
backend=biber,
style=authoryear,
autocite=inline
]{biblatex}
%\DeclareLanguageMapping{english}{english-apa}

@@

\section{Introduction}
\begin{frame}
\frametitle{Introduction}
\begin{itemize}
	\item Covering problems
	\begin{itemize}
		\item Minimum Cover Problem \autocite{karp}
		\item Maximal Covering Problem \autocite{church:1974}
	\end{itemize}

\end{itemize}

\begin{figure}
	\caption{Minimum Vertex Cover and its maximal counterpart. The colored edges are in the cover.}
	\input{figures/cover.tex}

	\source{Elaborated by the author.}
\end{figure}
\end{frame}

\begin{frame}{Introduction}
\begin{itemize}

	\item Maximal Covering Location Problem (MCLP)

	\begin{itemize}
		\item Introduce at first for networks. Facilities are placed on nodes, covering a radius of neighboring vertexes.
	\end{itemize}
	\item Planar Maximal Covering Location Problem (PMCLP)
	\begin{itemize}
		\item One disk is 3SUM-Hard \autocite{3SUM-kopelowitz:2014}.
		\item One disk: $\bigO(n^2)$ and $\bigO(n^2\log{n})$ algorithms.
		\item $m$ disks: $\bigO(n^{2m-1}\log{n})$ algorithm.
	\end{itemize}
	\item Goals
	\begin{itemize}
		\item Develop a $\bigO(n^2\log{n})$ algorithm for the one disk case.
		\item Adapt it for the $m$ ellipses case creating a $\bigO(n^{2m})$ algorithm.
	\end{itemize}
\end{itemize}

@@

@@

	$MCD(\Pp, 1)$ is the problem of placing one disk on the plane to cover a subset of a demand set $\Pp$, with $n$ points, maximizing the weights of the covered points.

@@

	\begin{itemize}
		\item $\Pp=\{p_1,\dots,p_n\}$ is the demand set with weights $w(p_i)>0$.
		\item $w(A)$, $A\subset \Pp$, is the sum of weights of the points in $A$.
		\item $D(q)$ is a unit disk with center at point $q$.
		%\item \autocite{chazelle:1986} proposed a $\bigO(n^2)$ algorithm
		%\item \autocite{drezner} proposed a $\bigO(n^2\log{n})$ which our work is based on
		%\item We will introduce an equivalent problem...
	\end{itemize}
\end{frame}

\begin{frame}{Maximal Covering by Disks}{One disk}
\begin{figure}
	\caption{An instance of $MCD(\Pp,1)$.}
	\includegraphics[scale=0.6]{figures/mcd_one_disk.pdf}
	\source{Elaborated by the author.}
\end{figure}
\end{frame}

\begin{frame}{Maximal Covering By Disks}{One disk}

	Works and results found in the literature:

	\begin{itemize}
		\item $MCD$ is as difficult the problem of given $n$ numbers, find three of them that sum to $0$. Proved by \autocite{aronov:2008}.

		\item In \autocite{drezner} a $\bigO(n^2\log{n})$ algorithm was developed. The idea of our algorithm to sort the intersections by their angles comes from here.

		\item In \autocite{chazelle:1986}, a $\bigO(n^2)$ algorithm was developed. It actually solves an equivalent problem which is introduced next.
	\end{itemize}

@@

	\begin{itemize}
		\item The disks are fixed with centers at $\Pp =\{p_1,\dots,p_n\}$ with $w_k=w(p_k)$.
		\item A clique is a non-empty intersection area of a subset of disks. We search for only a point in an optimal clique.
		\item An optimal solution for the maximum weight clique is an optimal solution for $MCD(\Pp,1)$.
	\end{itemize}
\end{frame}

\begin{frame}{Maximum Weight Clique Problem}
	\begin{figure}
		\only<1>{\caption{An instance of $MCD(\Pp, 1)$.}}
		\only<2>{\caption{An instance of the Maximum Weight Clique Problem obtained from an instance of $MCD(\Pp, 1)$.}}
		\only<3>{\caption{An instance of the Maximum Weight Clique Problem obtained from an instance of $MCD(\Pp, 1)$. In gray, the optimal solution.}}
		\only<1>{\includegraphics[scale=0.55]{figures/mwc_1.pdf}}
		\only<2>{\includegraphics[scale=0.55]{figures/mwc_2.pdf}}
		\only<3>{\includegraphics[scale=0.55]{figures/mwc_3.pdf}}
		%\only<2>{\input{figures/intro_b.tex}}
		\source{Elaborated by the author.}
	\end{figure}
\end{frame}

%\begin{frame}{Maximum Weight Clique Problem}{Equivalence}
%		\begin{figure}[H]
%		\centering

%		\only<1>{}

%		\only<2>{\caption{An instance of the Maximum Weight Clique Problem obtained from an instance of $MCD(\Pp, 1)$. In gray, the optimal solution.}}

%		\only<1>{\input{figures/max_w_clique_v2.tex}}
%		\only<2>{\input{figures/max_w_clique_b.tex}}
      %\center{\includegraphics[width=.7\textwidth]{figures/mcd_instance.png}}
%		\source{Elaborated by the author.}
%		\label{fig:mcd_instance}
%	\end{figure}
%\end{frame}

\begin{frame}{Maximum Weight Clique Problem}{Algorithm}
	Defining $\Gamma_+(i,j)$ and $\Gamma_-(i,j)$:\\~\\

	Let $D_i$ (at the origin) and $D_j$ be two unit disks that intersect at two points.

	\begin{itemize}
		\item We know that the two intersection points define two arcs in $D_i$.
		\item One of the arcs bounds $D_i\cap D_j$. That is the one we want to determine.

		\item We can determine the polar angles of the two intersection points.
		%\item Assume $D_i$ is at the origin.
		\item Assuming counter-clockwise direction, we define $\Gamma_+(i,j)$ and $\Gamma_-(i,j)$ as the angles of intersection that determines the arc of $D_i$ that bounds $D_i \cap D_j$.
	\end{itemize}
\end{frame}

\begin{figure}
	\caption{$\Gamma_+(1,2)$ and $\Gamma_-(1,2)$ example.}
	\includegraphics[scale=0.65]{figures/gammas.pdf}
		\source{Elaborated by the author.}
\end{figure}
\end{frame}

@@

	\begin{itemize}
		\item There are $\bigO(n^2)$ intersections among $n$ disks.

		\item Sorting takes $\bigO(n^2\log{n})$.

@@

		\item It can be implemented in $K\log{n}$ where $K$ is the number of intersections \autocite{bentley:1979}.
	\end{itemize}

@@

		with $\le$ representing the set of covered points.
	\end{block}

@@

	\begin{equation*}
	\max_q |\Pp \cap E(q)|,
	\end{equation*}

	\begin{itemize}
		\item $E(q)$ is an axis-parallel ellipse with center point $q$.
		\item Assuming unit weights for now.
		\item Same algorithm for one disk.
	\end{itemize}

@@

	\begin{itemize}
		\item $\E$ is a set of $m$ ellipses.
		\item \autocite{canbolat} is the very first study on the problem. Slow exact method, a heuristic one was proposed.
		\item \autocite{andreta} proposed a method that breaks the problem into smaller optimization ones. Also, they developed a method for the non-axis-parallel case.
	\end{itemize}
\end{frame}

\begin{frame}[fragile]{Maximal Covering by Ellipses}{Pre-processing that finds every possible coverage for ellipse $E_i$}
	\begin{algorithm}[H]
		%\caption{Algorithm for $MCE(\Pp, a, b)$ with unit weights}\label{algoritmo:mce1}
		\begin{algorithmic}[1]
			\STATE $A \gets \bigcup_{j \in I_i} \{\Gamma_+(i,j) \cup \Gamma_-(i,j)\}$
			\STATE $Z \gets \{\}$
			\STATE $Cov \gets \{p_i\}$
			\FOR{$cnt=1..2$} 
			\FOR{$a \in A$}
			\STATE Let $p_a$ be the point that intersects $E_i$ at angle $a$. 
			\IF{$a$ is a starting angle}
			\STATE $Cov \gets Cov \cup \{p_a\}$
			\ELSE
			\STATE $Cov \gets Cov \setminus \{p_a\}$
			\ENDIF
			\STATE $Z \gets Z \cup \{Cov\}$
			\ENDFOR
			\ENDFOR

@@

	\begin{itemize}
		\item The algorithm for $m$ ellipses tries every possible assignment of coverage for every one of the ellipses.
		\item Run-time complexity of $\bigO(n^{2m})$.

@@

		\item The unit-weight assumption can be easily dropped.
	\end{itemize}
